---
title: "Cell size data routine checks"
output: 
  html_document:
    toc: true
    toc_float: true
---

# Data import and setup

Libraries and data import and removing irrelevant data
```{r setup, include=FALSE}
knitr::opts_chunk$set(fig.width=6, fig.height=4, fig.path='Figs/sizes/',
                      echo = TRUE, message = FALSE, warning = FALSE)


library(tidyverse)
library(readxl)
library(ggridges)
library(car)
library(ggpubr)
library(stringi)
library(lme4)
library(cowplot)
library(scales)

# theme_set(theme_bw()+
#   theme(axis.text=element_text(size=16),
#   axis.title.x = element_text(size = 20),
#   axis.title.y = element_text(size = 20)))

theme_set(theme_bw()+
            theme(axis.text=element_text(size=6),
  axis.title.x = element_text(size = 8),
  axis.title.y = element_text(size = 8),
  strip.background=element_rect(fill="#F1F1F1"),
  strip.text = element_text(size=8)))


# theme_set(theme_bw()+
#             theme(axis.text=element_text(size=10),
#   axis.title.x = element_text(size = 12),
#   axis.title.y = element_text(size = 12),
#   strip.background=element_rect(fill="#F1F1F1"),
#   strip.text = element_text(size=12)))


select <- dplyr::select

jevergreen <- "#218D3A"

jever_3color <- c("#297C8C", "#116D2B", "#B59557")

jever_9color <- c("#297C8C", "#41C4DE", "#49DDFA",
                  "#116D2B","#1BAD44","#24E85B",
                  "#B59557", "#D1AC65", "#FACE79")

palette_9color <- c("lightsalmon4","lightsalmon3","lightsalmon2",
                    "turquoise4", "turquoise3", "turquoise2",
                    "seagreen4", "seagreen3", "seagreen2")

marrit_cells <- read_excel("Experiment 2 cell sizes 01.04.21_MJ.xlsx", 
    sheet = "all cell measurements") %>% 
  filter(entrant=='MJ') %>% 
  dplyr::select(plate:note2)
summary(marrit_cells)


svenja_cells <- read_excel("Experiment 2 cell sizes 20.03.21_SW_final.xlsx", 
    sheet = "all cell measurements") %>% 
  filter(entrant=='SW') %>% 
  dplyr::select(plate:note2)
summary(svenja_cells)


sizechecks <- bind_rows(marrit_cells, svenja_cells)
summary(sizechecks)

# summary reveals a min value of 0 for distance- looks like there is just one mistake here that should be removed
sizechecks %>% filter(distance<1) %>% print(n = Inf)

#sizechecks <- read_excel("C:/Users/pktho/OneDrive/Experiment 2 ALL SYSTEMATIC PHOTOS/Experiment 2 cell sizes 14.01.21_PTchanges.xlsx", 
#    sheet = "all cell measurements")

#sizechecks <- sizechecks[1:17]
sizechecks <- sizechecks %>% 
  filter(entrant!="PT") %>% 
  mutate(distance=distance/1000) %>% 
  select(-angle) %>% 
  mutate(species=ifelse(species=="thalasso", "tnema", species))
  #because it thinks commas are thousands and not decimal point but careful with this!!


summary(sizechecks)
str(sizechecks)

sizechecks %>% 
  group_by(plate, well, species) %>% 
  summarize(count = n())

```

>Checking that the dimensions make sense for each species. On second thought though I think this would be best to check in excel to make sure I change errors to the right dimension (like would 'l' for dity really be h or w?)

```{r eval=FALSE, include=FALSE}
#yeah for example there are some that have dimension labels that will cause problems
sizechecks %>% filter(species =='dity', dimension == 'l')

#the only way I can think to di this now is to separate by species
dity <- subset(sizechecks, species=="dity")
rhizo <- subset(sizechecks, species=="rhizo")
tnema <- subset(sizechecks, species=="tnema")

dity %>% filter(dimension == 'l')

dity <- dity %>% 
  mutate(dimension=ifelse(dimension=='l', 'h', dimension))

dity %>% filter(dimension == 'l')

##^^example of how I might do this but holding off on this for now**



```


Filtering and checking numbers of observations by treatments and then by each plate/well combination

**so far this is total obs not total number of cells so maybe move to after pivoting and mutating stuff**

```{r}

#checks where data is generally sparse or not

grouped <- sizechecks %>% group_by(nuts, temp, species_richness, strains) %>% 
  summarize(count=n())

ggplot(grouped, aes(strains, count))+
  geom_col(aes(fill=nuts), position = 'dodge')+
  facet_grid(temp~species_richness)+
  ylab('Total number of measurements')

#checks which wells have been done already

grouped2 <- sizechecks %>% group_by(plate, well, species_richness, strains, nuts, temp) %>% 
  summarize(count=n())
```

Check where the problems are so pivot_wider doesn't drive me crazy

```{r}
#makes an object of all the cell_ids that appear too many times

cell_id_checks <- sizechecks %>% 
  count(cell_id) %>% 
  filter(n>2)

FIX <- nrow(cell_id_checks)

biovolumes_wide <- sizechecks %>%
  pivot_wider(names_from = dimension,
              values_from = distance,
              values_fn = length, # THIS GIVES THE LENGTH, not actual values, just for checking data
              id_cols = cell_id)

#not sure why this gives a different number than the chunk above where it gives 56 with too many cell_ids, this gives 174??

```

> there are currently `r FIX` cell ID's that have too much data and need to be fixed

Using pivot wider to check if there are errors with cell_id (more than one or 2 per cell id = problematic)

```{r}

biovolumes_wide_check_counts <- biovolumes_wide %>% filter(h>1 | w>1 | d>1 | h>1)

```

> For now, attempting to exclude ones where the cell_ids are weird, but for the final analysis need to either manually fix those in excel, since that might be tough to do in R

> And even after I exclude the weird ones it still gives issues with pivot_wider :/

```{r}
#join the object with all the weird cell_ids and filter those out since those are the only nas, which gets ride of a couple hundred rows
biovolumes_wide_clean <- left_join(sizechecks, cell_id_checks) %>%
  filter(is.na(n))

biovolumes_wide_clean2 <- biovolumes_wide_clean %>%
  pivot_wider(names_from = dimension,
              values_from = distance, values_fn=length)

biovolumes_wide_check_counts <- biovolumes_wide_clean2 %>% filter(h>1 | w>1 | d>1 | h>1)

bad <- biovolumes_wide_check_counts[11] %>% 
  mutate(badrows='bad')

biovolumes_wide_clean3 <- left_join(sizechecks, bad) %>% 
  filter(is.na(badrows))

biovolumes_wide_clean4 <- biovolumes_wide_clean3 %>%
  select(cell_id, dimension, distance) %>% 
  pivot_wider(names_from = dimension,
              values_from = distance)

              
              
```

>Since pivot_wider is being really shitty, I'll just take the mean values for things that are weird, but this needs to be fixed. for preliminary work only about `r round(100*(FIX/nrow(sizechecks)), 2)`% are weird so this can be temporarily ignored

```{r}

biovolumes_wide <- sizechecks %>%
  pivot_wider(names_from = dimension,
              values_from = distance,
              values_fn = mean) #mean == wrong, so do it right later
              #id_cols = cell_id)

```


Separating the broken vs non-broken cells, separating species

```{r}
notbroken <- filter(biovolumes_wide, is.na(broken)) %>% 
  mutate(cell_broken = "n")
broken <- biovolumes_wide %>% filter(broken==1) %>% 
  mutate(cell_broken = "y")
```

Attempting to deal with the broken cells correctly

1. make objects for each species
2. calculate mean ratio of measured dimensions
3. use mutate to calculate unmeasured dimensions

```{r}
dity_intact <- subset(notbroken, species=="dity")
rhizo_intact <- subset(notbroken, species=="rhizo")
tnema_intact <- subset(notbroken, species=="tnema")

dity_mean_ratio <- mean(dity_intact$h/dity_intact$w, na.rm = TRUE)

rhizo_mean_ratio <- mean(rhizo_intact$h/rhizo_intact$d, na.rm=TRUE)

tnema_mean_ratio <- mean(tnema_intact$l/tnema_intact$h, na.rm = TRUE)

broken <- broken %>% 
  mutate(h = ifelse(species=='dity', w*dity_mean_ratio, h)) %>% 
  mutate(h = ifelse(species=='rhizo', d*rhizo_mean_ratio, h)) %>%
  mutate(l = ifelse(species=='tnema', h*tnema_mean_ratio, l))
  
biovolumes_with_broken <- bind_rows(dity_intact, rhizo_intact, tnema_intact, broken)

biovolumes_with_broken <- biovolumes_with_broken %>%
  mutate(biov='NA') %>% 
  mutate(biov = ifelse(species=='dity', w*w*h/2, biov)) %>% 
  mutate(biov = ifelse(species=='rhizo', (pi/4)*d^2*h, biov)) %>%
  mutate(biov = ifelse(species=='tnema', l*h*h, biov)) %>% 
  mutate(biov=as.integer(biov))

biovolumes_with_broken <- biovolumes_with_broken %>% 
  mutate(AR ='NA') %>%
  mutate(AR = ifelse(species=='dity', h/w, AR)) %>% 
  mutate(AR = ifelse(species=='rhizo', h/d, AR)) %>%
  mutate(AR = ifelse(species=='tnema', l/h, AR)) %>% 
  mutate(AR=as.numeric(AR))

```

```{r}
biovolumes_with_broken <- biovolumes_with_broken %>%
  mutate(fullname = "NA") %>% 
  mutate(fullname = ifelse(div_tx=='DB', 'Ditylum (Borkum)', fullname))%>%
  mutate(fullname = ifelse(div_tx=='DH', 'Ditylum (Helgoland)', fullname))%>% 
  mutate(fullname = ifelse(div_tx=='DK', 'Ditylum (Kiel)', fullname))%>% 
  mutate(fullname = ifelse(div_tx=='RB', 'Rhizosolenia (Borkum)', fullname))%>% 
  mutate(fullname = ifelse(div_tx=='RH', 'Rhizosolenia (Helgoland)', fullname))%>% 
  mutate(fullname = ifelse(div_tx=='RK', 'Rhizosolenia (Kiel)', fullname))%>% 
  mutate(fullname = ifelse(div_tx=='TB', 'Thalassionema (Borkum)', fullname))%>% 
  mutate(fullname = ifelse(div_tx=='TH', 'Thalassionema (Helgoland)', fullname))%>% 
  mutate(fullname = ifelse(div_tx=='TK', 'Thalassionema (Kiel)', fullname))%>% 
  mutate(species_full="NA") %>% 
  mutate(species_full = ifelse(species=='dity', 'Ditylum brightwelii', species_full))%>% 
  mutate(species_full = ifelse(species=='rhizo', 'Rhizosolenia setigera', species_full))%>% 
  mutate(species_full = ifelse(species=='tnema', 'Thalassionema nitzschiodes', species_full)) %>% 
  mutate(fullname=fct_rev(fullname)) %>% 
  mutate(nuts_temp=paste(nuts, temp, sep = "."))


# this is trying to do community weighted whatever, which I am giving up on

# biovolumes_with_broken <- biovolumes_with_broken %>% 
#   group_by(plate, well, species_richness, strains,
#            temp, nuts, nuts_temp) %>% 
#   mutate(sum_biov_per_well = sum(biov)) %>% # total biovolume that was measured in transects per well
#   mutate(pi_biov = biov/sum_biov_per_well) %>% #proportion of biovolume that each cell represents of what was measured/counted
#   mutate(xi_pi_biov = biov*pi_biov) %>%  # weight each cell by its proportion of the assessed biovolume
#   mutate(CWM_kindof = sum(xi_pi_biov),
#          CW_distance = pi_biov*(biov-CWM_kindof)^2)


summary(biovolumes_with_broken)
dim(biovolumes_with_broken)

```

See how aspect ratios are distributed for the ones that are actual measurements of unbroken cells - this would validate the approach of using average AR values to impute missing dimensions

```{r}

biovolumes_with_broken %>% 
  filter(cell_broken == "n", species_richness==1, strains==1) %>% 
  ggplot(aes(AR))+
  geom_histogram()+
  facet_wrap(~fullname, scales = "free")

```


```{r}
mono_biov <- filter(biovolumes_with_broken, species_richness==1, strains==1)
summary(mono_biov)


highdiv_biov <- filter(biovolumes_with_broken, species_richness==3, strains==3)

```



>setting up summary metrics by well, including CV that I did at first, and then actually trying a CWM and CW variance approach

```{r}

size_by_well <- biovolumes_with_broken %>% 
  group_by(plate, well, species_richness, strains,
           temp, nuts, nuts_temp, fullname) %>% 
  summarize(var_biov=var(biov),
            mean_biov=mean(biov),
            CV_biov=sd(biov)/mean(biov),
            var_AR = var(AR),
            mean_AR = mean(AR),
            CV_AR = sd(AR)/mean(AR)) %>% 
  mutate(plate=as.character(plate)) %>% 
  mutate(temp = as.character(temp)) %>% 
  filter(!is.na(CV_biov))

size_by_well$well<- stri_reverse(size_by_well$well)

write_csv(size_by_well, file = "size_by_well.csv")


# check how the biovolume weighted mean and variance look compared to the raw (abundance-weighted) moments

ggplot(size_by_well, aes(log(mean_biov), log(CWM_kindof)))+
  geom_point()


ggplot(size_by_well, aes(log(var_biov), log(CW_variance)))+
  geom_point()

ggplot(size_by_well, aes(var_biov, CW_variance))+
  geom_point()+
  scale_x_log10()+
  scale_y_log10()


#splits <- strsplit(size_by_well$well, "")[[1]]
#size_by_well$well <-  intToUtf8(rev(utf8ToInt(size_by_well$well)))

#mutate(plate_well = paste(plate, well)) %>% 

size_by_well$nuts_temp <- ordered(size_by_well$nuts_temp, levels = (c("low.8","low.12","low.16","high.8", "high.12", "high.16")))

#allmaxRFUandfits <- read_csv("allmaxRFUandfits17.7.20_withK400allowed.csv")

#sizesandRFU <- left_join(allmaxRFUandfits, size_by_well) %>% 
#  mutate(totalstrains=species_richness*strains)

```


```{r}
size_by_well %>% count(species, strains, temp) %>% print(n = Inf)
```

# Effects of environment on cell sizes and shapes

## monocultures and sizes
```{r}
mono_biov$nuts_temp <- ordered(mono_biov$nuts_temp, levels = (c("low.8","low.12","low.16",
                                                            "high.8", "high.12", "high.16")))
highdiv_biov$nuts_temp <- ordered(highdiv_biov$nuts_temp, levels = (c("low.8","low.12","low.16",
                                                            "high.8", "high.12", "high.16")))
```

# >> nice fig1 option <<

```{r}

mono_biov <- mono_biov %>% 
  mutate(origin = "NA") %>% 
  mutate(origin = ifelse(grepl("Borkum", fullname),"Borkum", origin)) %>% 
  mutate(origin = ifelse(grepl("Helgoland", fullname), "Helgoland", origin)) %>% 
  mutate(origin = ifelse(grepl("Kiel", fullname), "Kiel", origin))


fig1a <- mono_biov %>% 
  filter(cell_broken=="n") %>% 
  ggplot(aes(x = biov, y = fullname))+
  geom_density_ridges(aes(fill=species_full), 
                      #size = 1.2,
                      size = 0.6)+
  scale_x_log10(breaks = trans_breaks("log10", function(x) 10^x),
              labels = trans_format("log10", math_format(10^.x))) +
  ylab('Diatom strain identity')+
  xlab(bquote("Cell volume" ~(µm^3)))+
  #scale_fill_brewer(palette = "Paired")+
  scale_fill_manual(values = jever_3color)+
  #scale_fill_manual(values = palette_9color)+
  theme(legend.position = "none")
fig1a
ggsave("density_plots.png",  width = 150, height = 100, units = "mm", scale = 1, dpi = 600, device = "png")

fig1b <- mono_biov %>% 
  filter(cell_broken=="n") %>% 
  ggplot(aes(x = AR, y = fullname))+
  geom_density_ridges(aes(fill=species_full), size = 0.6, scale = 1)+
  scale_x_log10()+
  ylab('Diatom strain identity')+
  xlab("Cell shape (aspect ratio)")+
  #scale_fill_brewer(palette = "Paired")+
  scale_fill_manual(values = jever_3color)+
  theme(legend.position = "none",
        # axis.title.y = element_blank(), #use these 3 lines to put a and b on top
        # axis.text.y = element_blank(),
        # axis.ticks.y = element_blank()
        )
fig1b

fig1c <- mono_biov %>% 
  filter(cell_broken=="n") %>% 
  #filter(!is.na(biov), !is.na(AR), is.na(broken)) %>% 
  mutate(fullname = fct_rev(fullname)) %>% 
  ggplot(aes(biov, AR, shape = origin, fill=species_full))+
  #scale_x_discrete(limits = rev)+
  #scale_fill_brewer(palette = "Dark2", name = "Strain identity")+
  #scale_fill_manual(values = jever_3color, name = "Species")+
  scale_fill_manual(values = jever_3color, guide = "none")+
  geom_point(color = "black", size=1.5, alpha = 0.6)+
  scale_shape_manual(values = c(21, 22, 24), name = "Strain origin")+
  scale_x_log10(breaks = trans_breaks("log10", function(x) 10^x),
              labels = trans_format("log10", math_format(10^.x))) +
  scale_y_log10()+
  theme(legend.position = "none")+
  # theme(legend.position = c(0.8, 0.8),
  #       legend.text = element_text(size = 6),
  #       legend.key = element_blank())+
  ylab("Cell shape (aspect ratio)")+
  xlab(bquote("Cell volume" ~(µm^3)))#
fig1c
  

left_side <- plot_grid(fig1a, fig1b, ncol=1, nrow = 2, labels = c('A', 'B'))

plot_grid(left_side, fig1c, labels = c('', 'C'), rel_widths = c(1, 1.5))

ggsave("July2023figs/fig1_sizesxshapes.png",  width = 180, height = 120, units = "mm", scale = 1, dpi = 600, device = "png")


# another option to shift things around
left_side2 <- plot_grid(fig1a, fig1b, ncol=2, nrow = 1, labels = c('A', 'B'))

plot_grid(left_side2, fig1c, ncol = 1, nrow = 2, labels = c('', 'C'), rel_heights  = c(0.5, 1))

ggsave("July2023figs/fig1_sizesxshapes_v2.png",  width = 180, height = 180, units = "mm", scale = 1, dpi = 600, device = "png")

```


```{r}
ggplot(mono_biov, aes(x = biov, y = fullname))+
  geom_density_ridges(aes(fill=fullname), size = 1.2)+
  scale_x_log10()+
  ylab('Diatom strain identity')+
  xlab(bquote("Biovolume" ~(µm^3)))+
  theme(legend.position = "none")+
  geom_hline(yintercept =3.5, lty='dashed')
ggsave("mono_sizes2.png")

ggplot(mono_biov, aes(x = biov, y = as.factor(nuts_temp)))+
  geom_density_ridges(aes(fill=species_full), size = 1.2)+
  facet_wrap(~fct_rev(fullname), scales = 'free')+
  scale_x_log10()+
  ylab('Nutrient × temperature environment')+
  xlab(bquote("Biovolume" ~(µm^3)))+
  theme(legend.position = "none")


ggplot(biovolumes_with_broken, aes(x = biov, y = nuts_temp))+
  geom_density_ridges()+
  #scale_x_log10()+
  #ylab('Diatom strain identity')+
  xlab(bquote("Biovolume" ~(µm^3)))+
  #theme(legend.position = "none")+
  facet_wrap(species_richness~strains)+
  ggtitle('trying to show how nutsxtemp has different effects on size for different
          diversity levels but maybe not possible this way....')


```

##monocultures and shapes

```{r}

ggplot(mono_biov, aes(x = AR, y = fullname))+
  geom_density_ridges(aes(fill=species_full), size = 1.2)+
  scale_x_log10()+
  ylab('Diatom strain identity')+
  xlab("Aspect ratio")+
  theme(legend.position = "none")
  

ggplot(mono_biov, aes(x = AR, y = as.factor(nuts_temp)))+
  geom_density_ridges(aes(fill=species_full), size = 1.2)+
  facet_wrap(~fct_rev(fullname), scales = 'free')+
  scale_x_log10()+
  ylab('Nutrient × temperature environment')+
  xlab("Aspect ratio")+
  theme(legend.position = "none")


```


## 3 species 3 strains

Density plots of size by temperature for high diversity treatments

```{r}

########

ggplot(highdiv_biov, aes(x = biov, y = as.factor(nuts_temp)))+
  geom_density_ridges(fill='#00AFBB', size = 1.4)+
  scale_x_log10()+
  ylab('Nutrient × temperature environment')+
  xlab(bquote("Biovolume" ~(µm^3)))+
  theme(legend.position = "none")

########



ggplot(highdiv_biov, aes(as.factor(temp), biov,  fill=nuts))+
  geom_boxplot()+
  scale_y_log10()

ggplot(highdiv_biov, aes(as.factor(temp), biov,  fill=nuts))+
  geom_violin()+
  scale_y_log10()

```

## All diversity treatments

```{r}
# ggplot(biovolumes_with_broken, aes(x = biov, y = as.factor(temp)))+
#   geom_density_ridges()+
#   scale_x_log10()
# 
# ggplot(biovolumes_with_broken, aes(x = biov, y = as.factor(nuts)))+
#   geom_density_ridges()+
#   scale_x_log10()
# 
# ggplot(biovolumes_with_broken, aes(biov))+
#   geom_density()+
#   facet_grid(temp~nuts)+
#   scale_x_log10()
```

# Effects of species and strain richness on size diversity


> not having any fucking luck with left_join today so this shit doesn't work

>FIXED THE STUPID BACKWARDS WELL NAMES!!!!!!!!!!!!!!!!!!!!!!!! then left_join worked



```{r}



```

## plots
 
```{r}
# add this for nonlinear fit
# stat_smooth(method = "lm", formula = y ~ x + I(x^2))+

A <- ggplot(size_by_well, aes(species_richness, CV_biov))+
  geom_point()+
  #geom_smooth(method="lm")+
  stat_smooth(method = "lm")+
  xlab("Species richness")+
  ylab("Cell size variation (CV)")+
  scale_x_continuous(breaks=c(1,2,3))+
  scale_y_log10()
A

B <- ggplot(size_by_well, aes(strains, CV_biov))+
  geom_point()+
  #geom_smooth(method="lm")+
  stat_smooth(method = "lm")+
  xlab("Strain richness per species")+
  ylab("Cell size variation (CV)")+
  scale_x_continuous(breaks=c(1,2,3))+
  scale_y_log10()+
  theme(axis.title.y=element_blank(),
        axis.text.y=element_blank(),
        axis.ticks.y=element_blank())
B

ggarrange(A,B)
ggsave("strain_and_sp_rich_vs_CVsize.png")


A <- ggplot(size_by_well, aes(species_richness, var_biov))+
  geom_point()+
  #geom_smooth(method="lm")+
  stat_smooth(method = "lm")+
  xlab("Species richness")+
  ylab("Cell size variance")+
  scale_x_continuous(breaks=c(1,2,3))+
  scale_y_log10()
A

B <- ggplot(size_by_well, aes(strains, var_biov))+
  geom_point()+
  #geom_smooth(method="lm")+
  stat_smooth(method = "lm")+
  xlab("Strain richness per species")+
  ylab("Cell size variance")+
  scale_x_continuous(breaks=c(1,2,3))+
  scale_y_log10()
B

ggarrange(A,B)


#### USE THIS - SEEMS BEST APPROACH ####

A <- ggplot(size_by_well, aes(species_richness, CW_variance))+
  geom_point()+
  #geom_smooth(method="lm")+
  stat_smooth(method = "lm", color = jevergreen, size = 2)+
  xlab("Species richness")+
  ylab("Community-weighted size variance")+
  scale_x_continuous(breaks=c(1,2,3))+
  scale_y_log10()
A


B <- ggplot(size_by_well, aes(strains, CW_variance))+
  geom_point()+
  #geom_smooth(method="lm")+
  stat_smooth(method = "lm", color = jevergreen, size = 2)+
  xlab("Strain richness of each species")+
  ylab("Community-weighted size variance")+
  scale_x_continuous(breaks=c(1,2,3))+
  scale_y_log10()
# 
#   theme(axis.title.y=element_blank(),
#         axis.text.y=element_blank(),
#         axis.ticks.y=element_blank())
B

p <- ggarrange(A,B)
p

ggsave("richness_vs_CWvariance.png")


library(cowplot)
plot_grid(A, B)

fit <- lm(data= size_by_well, log(CW_variance)~species_richness*strains)
summary(fit)
plot(fit)
avPlots(fit)

####




A <- ggplot(size_by_well, aes(species_richness, CWM_kindof))+
  geom_point()+
  #geom_smooth(method="lm")+
  stat_smooth(method = "lm")+
  xlab("Species richness")+
  ylab("Community-weighted mean size")+
  scale_x_continuous(breaks=c(1,2,3))+
  scale_y_log10()
A

B <- ggplot(size_by_well, aes(strains, CWM_kindof))+
  geom_point()+
  #geom_smooth(method="lm")+
  stat_smooth(method = "lm")+
  xlab("Strain richness of each species")+
  scale_x_continuous(breaks=c(1,2,3))+
  scale_y_log10()+
  theme(axis.title.y=element_blank(),
        axis.text.y=element_blank(),
        axis.ticks.y=element_blank())
B

p <- ggarrange(A,B)
annotate_figure(p, "Huh, same trends for CWM as CW variance by sp and strain richness")

A <- ggplot(size_by_well, aes(species_richness, mean_biov))+
  geom_point()+
  #geom_smooth(method="lm")+
  stat_smooth(method = "lm")+
  xlab("Species richness")+
  ylab("mean size just raw abundance weighted")+
  scale_x_continuous(breaks=c(1,2,3))+
  scale_y_log10()
A

B <- ggplot(size_by_well, aes(strains, mean_biov))+
  geom_point()+
  #geom_smooth(method="lm")+
  stat_smooth(method = "lm")+
  xlab("Strain richness of each species")+
  scale_x_continuous(breaks=c(1,2,3))+
  scale_y_log10()+
  theme(axis.title.y=element_blank(),
        axis.text.y=element_blank(),
        axis.ticks.y=element_blank())
B

p <- ggarrange(A,B)
annotate_figure(p, "but using the raw (abundance-weighed) mean shows totally different trends")


ggplot(size_by_well, aes(species_richness, CV_biov))+
  geom_jitter(width = 0.02, alpha = 0.5,aes(color=as.factor(strains)))+
  geom_smooth(size = 1.5, se = FALSE, method = 'lm', aes(color = as.factor(strains)))+
  scale_y_log10()+
  facet_wrap(~nuts_temp, scales = 'free')+
  xlab("Species richness")+
  labs(color = "Strain richness")+
  scale_x_continuous(breaks=c(1,2,3))+
  ylab('Cell size variation (CV)')

ggplot(size_by_well, aes(strains, CV_biov))+
  geom_jitter(width = 0.02, alpha = 0.5,aes(color=as.factor(strains)))+
  geom_smooth(size = 1.5, se = FALSE, method = 'lm', aes(color = as.factor(species_richness)))+
  scale_y_log10()+
  facet_wrap(~nuts_temp, scales = 'free')+
  xlab("Strain richness")+
  labs(color = "Species richness")+
  scale_x_continuous(breaks=c(1,2,3))+
  ylab('Cell size variation (CV)')







# 
# 
# ggplot(size_by_well, aes(nuts_temp, CV_biov))+
#   geom_boxplot()+
#   #geom_smooth(method="lm")+
#   xlab("Nutrient x temperature environment")+
#   ylab("Cell size variation (CV)")+
#   scale_y_log10()
#
```

## size div vs biomass plots

```{r}

ggplot(full_df, aes(CV_biov, maxRFU))+
  geom_point()+
  #geom_smooth(method="lm")+
  stat_smooth(method = "lm")+
  xlab("Cell size variation (CV)")+
  ylab("Biomass (max fluorescence)")+
  scale_x_log10()+
  scale_y_log10()

C <- ggplot(full_df, aes(mean_biov, maxRFU))+
  geom_point()+
  #geom_smooth(method="lm")+
  stat_smooth(method = "lm")+
  xlab("Mean biovolume")+
  ylab("Biomass (max fluorescence)")+
  scale_x_log10()+
  scale_y_log10()

D <- ggplot(full_df, aes(var_biov, maxRFU))+
  geom_point()+
  #geom_smooth(method="lm")+
  stat_smooth(method = "lm")+
  xlab("Variance biovolume")+
  ylab("Biomass (max fluorescence)")+
  scale_x_log10()+
  scale_y_log10()

ggarrange(C,D)

C <- ggplot(full_df, aes(CWM_kindof, maxRFU))+
  geom_point()+
  #geom_smooth(method="lm")+
  stat_smooth(method = "lm")+
  xlab("Mean biovolume")+
  ylab("Biomass (max fluorescence)")+
  scale_x_log10()+
  scale_y_log10()

D <- ggplot(full_df, aes(CW_variance, maxRFU))+
  geom_point()+
  #geom_smooth(method="lm")+
  stat_smooth(method = "lm")+
  xlab("Variance biovolume")+
  ylab("Biomass (max fluorescence)")+
  scale_x_log10()+
  scale_y_log10()

ggarrange(C,D)

```

## models


> models for CV size with varying interactions etc

```{r}
fit1 <- lm(data=size_by_well, CV_biov~species_richness*strains)
summary(fit1)
avPlots(fit1)

fit2 <- lm(data=size_by_well, CV_biov~species_richness*strains + temp + nuts)
summary(fit2)
avPlots(fit2)

fit2.1 <- lm(data=size_by_well, CV_biov~species_richness*strains + temp*nuts)
summary(fit2.1)
avPlots(fit2.1)


fit3 <- lm(data=size_by_well, CV_biov~species_richness + strains + temp + nuts)
summary(fit3)
avPlots(fit3)


# just a reminder that I really shouldn't use nuts unless I measure a lot more cells there to balance things out....
size_by_well %>%
  group_by (nuts, temp, species_richness, strains) %>%
  summarize (n=length(pi_check)) %>% print(n = Inf)

###USE THIS FOR I THINK (UNLESS A MIXED EFFECT MODEL ACTUALLY WORKS AND/OR low nuts data grows)####
fit_CWvariance <- lm(data = filter(size_by_well, nuts=="high"), 
                                   log(CW_variance)~species_richness*strains + temp)
summary(fit_CWvariance)
avPlots(fit_CWvariance)
vif(fit_CWvariance)
plot(fit_CWvariance)

####



fit_CWvariance2 <- lm(data=size_by_well, log(CW_variance)~species_richness*strains)
summary(fit_CWvariance2)
avPlots(fit_CWvariance2)
vif(fit_CWvariance2)
plot(fit_CWvariance2)


# I don't understand why this is singular if I include intxn but works if additive...
fit_CWvariance_randomenvt <- lmer(data=filter(size_by_well, nuts=='high'), log(CW_variance)~species_richness + strains + (1|temp))
summary(fit_CWvariance_randomenvt)

# or at least try something along these lines with lme4...

#fit3 <- lmer(data=size_by_well, CV_biov ~ species_richness + strains + 1|temp)
#summary(fit3)


AIC(fit1, fit2, fit2.1, fit3, fit_CWvariance, fit_CWvariance2, fit_CWvariance_randomenvt)

```


> models for variance in size (otherwise same as above)

```{r}
fit1 <- lm(data=size_by_well, var_biov~species_richness*strains)
summary(fit1)
avPlots(fit1)

fit2 <- lm(data=size_by_well, var_biov~species_richness*strains + temp + nuts)
summary(fit2)
avPlots(fit2)

fit2.1 <- lm(data=size_by_well, var_biov~species_richness*strains + temp*nuts)
summary(fit2.1)
avPlots(fit2.1)


fit3 <- lm(data=size_by_well, var_biov~species_richness + strains + temp + nuts)
summary(fit3)
avPlots(fit3)



# or at least try something along these lines with lme4...

#fit3 <- lmer(data=size_by_well, CV_biov ~ species_richness + strains + 1|temp)
#summary(fit3)


AIC(fit1, fit2, fit2.1, fit3)

```



# Calculation of biovolume per ml

```{r}
```

# Shapes

```{r}

ggplot(mono_biov, aes(x = AR, y = div_tx))+
  geom_density_ridges()+
  scale_x_log10()
  
ggplot(mono_biov, aes(x = AR, y = as.factor(temp)))+
  geom_density_ridges()+
  facet_wrap(~div_tx)+
  scale_x_log10()

ggplot(mono_biov, aes(x = AR, y = as.factor(nuts)))+
  geom_density_ridges()+
  facet_wrap(~div_tx)+
  scale_x_log10()

```

## All diversity treatments

```{r}
ggplot(biovolumes_with_broken, aes(x = AR, y = as.factor(temp)))+
  geom_density_ridges()+
  scale_x_log10()

ggplot(biovolumes_with_broken, aes(x = AR, y = as.factor(nuts)))+
  geom_density_ridges()+
  scale_x_log10()

ggplot(biovolumes_with_broken, aes(AR))+
  geom_density()+
  facet_grid(temp~nuts)+
  scale_x_log10()
```

## biov vs AR for monos

```{r}
mono_biov %>% 
  filter(!is.na(biov), !is.na(AR), is.na(broken)) %>% 
  ggplot(aes(biov, AR, color=species))+
  geom_point()+
  scale_x_log10()+
  scale_y_log10()

mono_biov %>% 
  filter(!is.na(biov), !is.na(AR), is.na(broken)) %>% 
  ggplot(aes(biov, AR, shape = species_full, color=fct_rev(fullname)))+
  #scale_x_discrete(limits = rev)+
  geom_point(size=2.4)+
  scale_x_log10()+
  scale_y_log10()+
  ylab('Aspect ratio')+
  xlab(bquote("Biovolume" ~(µm^3)))+
  scale_color_discrete(name = "Strain identity")+
  scale_shape_discrete((name = "Species"))
ggsave("biov_x_AR.monos.png")

```

for ppt

```{r}

mono_biov %>% 
  filter(is.na(broken)) %>% 
  ggplot(aes(x = biov, y = div_tx))+
  geom_density_ridges()+
  scale_x_log10()+
  xlab("biovolume")+
  ylab("strain")


mono_biov %>% 
  filter(is.na(broken)) %>% 
  ggplot(aes(x = AR, y = div_tx))+
  geom_density_ridges()+
  scale_x_log10()+
  xlab("aspect ratio")+
  ylab("strain")

mono_biov %>% 
  filter(is.na(broken)) %>% 
  ggplot(aes(biov, AR, color=div_tx))+
  geom_point(size=2)+
  scale_x_log10()+
  scale_y_log10()+
  xlab("biovolume")+
  ylab("aspect ratio")


```


checking min and max sizes by strains because I just got 18s results back and they don't quite match with the actual sizes esp tnema

```{r}
mono_biov

mono_biov %>%
  filter(!is.na(h), !is.na(l)) %>% 
  group_by(div_tx) %>% 
  summarize(minh=min(h), minl=min(l), maxh=max(h), maxl=max(l))

sdfdsf <- mono_biov %>%
  group_by(div_tx) %>% 
  summarize(minh=min(h), minl=min(l), minw=min(w), mind=min(d))
            
            
mono_biov %>% 
  group_by(div_tx) %>%
  summarize(maxh=max(h), maxl=max(l), maxw=max(w), maxd=max(d))


```




```{r}
knitr::knit_exit()
```

